Parallel Information Set Generation for Kriegspiel

### 2012-06-27 sbac/paper.tex
spell check

\begin{abstract}
Information Set Generation is the identification of the set of paths in an imperfect information game tree that are
consistent with a player's observations.  The ability to reason about the possible game history is critical to the
performance of game playing agents.  In this work, we address the parallelization of information set generation in the context
of kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search
engine and discuss its performance under different load balancing strategies and with different grainsize parameters.
We achieve speedups of over 500 on 1024 processors, far exceeding previous scalability results for game tree search applications.
\end{abstract}

@@

Information Set Generation (ISG) is an operation that game tree search agents
can use to identify nodes in the game tree that are consistent with their
observations~\cite{richards12reasoning}.  Because the operation is relatively
expensive and may be repeatedly applied by a decision-making agent, its
efficiency is a key factor in an agent's overall performance.  In this work, we
show that ISG for kriegspiel can be efficiently parallelized.

@@

In the game of kriegspiel, information set generation is much more complicated.  The player knows the opponent's position with
certainty when the game begins, but after the initial move there may be varying levels of uncertainty about the location
of the opponent's pieces.  Unlike poker, it is not possible to simply permute all of the opponent's possible pieces over
all of the squares not occupied by the player's own pieces.  A configuration of pieces for the opponent is only valid if
it can be reached by a legal sequence of moves.  In the general case, finding the nodes in an information set is a
combinatorial search problem.

Information set generation is not the same problem as solving a game tree, although it is arguably a necessary
subroutine for quality game play in imperfect information games.  In this work, we describe an approach to generating
information sets for the game of kriegspiel.

@@

\Section{Performance Results}\label{Results}
In order to understand the effect of various parameters such as grain size and load balancing, we first evaluated the scalability of our implementation using a randomly selected problem instance. Next, we show that our implementation performs equally well for instances from actual games.
\subsection{Analysis using Random Instance}
The performance results reported in
this section were obtained by running our program on Bleuprint with up to 1024 processors. TODO: Blueprint H/w or citation

@@

\subsubsection{Iso-efficiency}
\begin{figure}[h]
\centering
\begin{minipage}{0.5\linewidth}
\includegraphics[width=\columnwidth]{plots/3depths.pdf}
\end{minipage}
\centering
\begin{minipage}{0.5\linewidth}
\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
\end{minipage}
\vspace{-0.1in}
\caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
(Bottom) Same data with a linear y-axis for the running time (to help see the isoefficiency.)}
\vspace{-0.2in}
\label{3depths}
\end{figure}

We next attempted to measure the isoefficiency [TODO:cite] of the problem---how is efficiency of the parallelization of the
algorithm affected by {\em size} of the problem?  To test this, we varied the number of plies in the test instance from
8 to 12. We used the work stealing load balancing strategy.  The results are shown in Figure~\ref{3depths}.   In the easiest case (depth 8), there isn't
enough additional parallel work to justify the jump from 256 to 1024 processors, and performance actually degrades.
Other than that, however, we see consistent speedups as we increase the number of processors.  The plot on the right helps to
illustrate the good isoefficiency of the problem.  As the size of the problem increases, the amount of parallelizable
work available also increases, and therefore the steepness of the speed curve is higher for larger problem sizes, as the
number of processors increases.

@@

Note also that information set generation is just one subroutine needed to reason about partially observable games.  Once plausible game histories have been found, an agent may simulate some sequences of moves up to previous moves made by opponents and then recursively call the information set generation procedure from the perspective of the opponent.  This results in many parallel applications of information set generation. Furthermore, given a node in an informaion set, it is common for an agent to perform forward search on that node (e.g., using MCTS).  With these two additional levels of parallelism on top of the information set generation problem, it is conceivable that the problem of playing partially observable games could scale efficiently to millions of processors or more.
\begin{table}[tphb]
\centering
\begin{tabular}{lrrc}
{\bf Authors} & {\bf Game} & {\bf Year} & {\bf Best Speedup (cores)} \\
Ferguson \& Korf & Othello & 1988 & 12(32) \\
Felton \& Otto & Chess & 1988 & 100(256) \\
Richards \etal & Kriegspiel & 2012 & 178(256) \\
Richards \etal & Kriegspiel & 2012 & {\bf 536(1024)} \\
\end{tabular}
\caption{Best Speedups reported for game tree applications.}
\vspace{-0.2in}
\label{bestspeedups}
\end{table}

@@

%Russell and Wolfe have performed some analyses on kriegspiel positions for which it possible to prove that one player
%has a forced win (that is, regardless of the configuration of the opponent's unseen pieces, the player has a sequence of
%moves that is guaranteed to produce a checkmate)~\cite{russell05efficient, wolfe07exploiting}.  The authors assume
%knowledge of the player's belief state.  They claim that under certain "aggressive" styles of play, it is uncommon for
%information sets to exceed 10,000 nodes in size.  They throw out (i.e., do not analyze) any positions that exceed this
%threshold.

>>>>

\begin{abstract}
Information Set Generation is the identification of the set of paths in an imperfect information game tree that are
consistent with a player's observations.  The ability to reason about the possible game history is critical to the
performance of game playing agents.  In this work, we address the parallelization of information set generation in the context
of Kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search
engine and discuss its performance under different load balancing strategies and with different grain-size parameters.
We achieve speedups of over 500 on 1024 processors, far exceeding previous scalability results for game tree search applications.
\end{abstract}

@@

Information Set Generation (ISG) is an operation that game tree search agents
can use to identify nodes in the game tree that are consistent with their
observations~\cite{richards12reasoning}.  Because the operation is relatively
expensive and may be repeatedly applied by a decision-making agent, its
efficiency is a key factor in an agent's overall performance.  In this work, we
show that ISG for Kriegspiel can be efficiently parallelized.

@@

In the game of Kriegspiel, information set generation is much more complicated.  The player knows the opponent's position with
certainty when the game begins, but after the initial move there may be varying levels of uncertainty about the location
of the opponent's pieces.  Unlike poker, it is not possible to simply permute all of the opponent's possible pieces over
all of the squares not occupied by the player's own pieces.  A configuration of pieces for the opponent is only valid if
it can be reached by a legal sequence of moves.  In the general case, finding the nodes in an information set is a
combinatorial search problem.

Information set generation is not the same problem as solving a game tree, although it is arguably a necessary
subroutine for quality game play in imperfect information games.  In this work, we describe an approach to generating
information sets for the game of Kriegspiel.

@@

\Section{Performance Results}\label{Results}
In order to understand the effect of various parameters such as grain size and load balancing, we first evaluated the scalability of our implementation using a randomly selected problem instance. Next, we show that our implementation performs equally well for instances from actual games.
\subsection{Analysis using Random Instance}
The performance results reported in
this section were obtained by running our program on Blueprint with up to 1024 processors. TODO: Blueprint H/w or citation

@@

\subsubsection{Iso-efficiency}
\begin{figure}[h]
\centering
\begin{minipage}{0.5\linewidth}
\includegraphics[width=\columnwidth]{plots/3depths.pdf}
\end{minipage}
\centering
\begin{minipage}{0.5\linewidth}
\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
\end{minipage}
\vspace{-0.1in}
\caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
(Bottom) Same data with a linear y-axis for the running time (to help see the Iso-efficiency.)}
\vspace{-0.2in}
\label{3depths}
\end{figure}

We next attempted to measure the Iso-efficiency [TODO:cite] of the problem---how is efficiency of the parallelization of the
algorithm affected by {\em size} of the problem?  To test this, we varied the number of plies in the test instance from
8 to 12. We used the work stealing load balancing strategy.  The results are shown in Figure~\ref{3depths}.   In the easiest case (depth 8), there isn't
enough additional parallel work to justify the jump from 256 to 1024 processors, and performance actually degrades.
Other than that, however, we see consistent speedups as we increase the number of processors.  The plot on the right helps to
illustrate the good isoefficiency of the problem.  As the size of the problem increases, the amount of parallelizable
work available also increases, and therefore the steepness of the speed curve is higher for larger problem sizes, as the
number of processors increases.

@@

Note also that information set generation is just one subroutine needed to reason about partially observable games.  Once plausible game histories have been found, an agent may simulate some sequences of moves up to previous moves made by opponents and then recursively call the information set generation procedure from the perspective of the opponent.  This results in many parallel applications of information set generation. Furthermore, given a node in an information set, it is common for an agent to perform forward search on that node (e.g., using MCTS).  With these two additional levels of parallelism on top of the information set generation problem, it is conceivable that the problem of playing partially observable games could scale efficiently to millions of processors or more.
\begin{table}[tphb]
\centering
\begin{tabular}{lrrc}
{\bf Authors} & {\bf Game} & {\bf Year} & {\bf Best Speedup (cores)} \\
Ferguson \& Korf & Othello & 1988 & 12(32) \\
Felton \& Otto & Chess & 1988 & 100(256) \\
Richards \etal & Kriegspiel & 2012 & 178(256) \\
Richards \etal & Kriegspiel & 2012 & {\bf 536(1024)} \\
\end{tabular}
\caption{Best Speedups reported for game tree applications.}
\vspace{-0.2in}
\label{bestspeedups}
\end{table}

@@

%Russell and Wolfe have performed some analysis on kriegspiel positions for which it possible to prove that one player
%has a forced win (that is, regardless of the configuration of the opponent's unseen pieces, the player has a sequence of
%moves that is guaranteed to produce a checkmate)~\cite{russell05efficient, wolfe07exploiting}.  The authors assume
%knowledge of the player's belief state.  They claim that under certain "aggressive" styles of play, it is uncommon for
%information sets to exceed 10,000 nodes in size.  They throw out (i.e., do not analyze) any positions that exceed this
%threshold.
